---
title: "PM2.5 Delhi, data sources"
author: "Maëlle Salmon"
date: "23 janvier 2016"
output: html_document
---

Maybe it's nicer to have a html document with the analysis!

So, the goal is to compare historic PM2.5 values for Delhi as found on the CPCB website by Eric Dodge to values queried from OpenAQ.

# Load packages

```{r, warning=FALSE, message=FALSE}
library("readr")
library("lubridate")
library("dplyr")
library("Ropenaq")
library("ggplot2")
```

# Check available locations for Delhi on OpenAQ

```{r}
Ropenaq::locations(city="Delhi", parameter="pm25")
# we'll use only the 4 first ones since the first one
# is US embassy data
locationsDelhi <- Ropenaq::locations(city="Delhi", 
                                     parameter="pm25")[1:4,]
```

# Load the CPCB historic data

```{r}
dataCPCB <- readr::read_csv("cpcb_ambient_panel.csv")
# change this name for compatibility with Open AQ name
dataCPCB$station[dataCPCB$station=="R K Puram"] <- "RK Puram"
# filter the locations we have with OpenAQ
dataCPCB <- dplyr::filter(dataCPCB,
                          station %in% locationsDelhi$location)
# now off to translating date
# I am too lazy for finding something more elegant
dataCPCB$dt_clean <- gsub("apr", "-04-", dataCPCB$dt_clean)
dataCPCB$dt_clean <- gsub("may", "-05-", dataCPCB$dt_clean)
dataCPCB$dt_clean <- gsub("jun", "-06-", dataCPCB$dt_clean)
dataCPCB$dt_clean <- gsub("jul", "-07-", dataCPCB$dt_clean)
dataCPCB$dt_clean <- gsub("aug", "-08-", dataCPCB$dt_clean)
dataCPCB$dt_clean <- gsub("sep", "-09-", dataCPCB$dt_clean)
dataCPCB$dt_clean <- gsub("oct", "-10-", dataCPCB$dt_clean)
dataCPCB$dt_clean <- gsub("nov", "-11-", dataCPCB$dt_clean)
dataCPCB$dt_clean <- gsub("dec", "-12-", dataCPCB$dt_clean)
dataCPCB <- dplyr::mutate(dataCPCB,
                          dateLocal=lubridate::dmy_hms(dt_clean))
# name the column differently
dataCPCB <- dplyr::mutate(dataCPCB,
                          historicValue=reading_value)
# drop useless columns
dataCPCB <- dplyr::select(dataCPCB,
                          - dt_clean,
                          - date_r,
                          - monitor_read,
                          - reading_value)
```

# Get Open AQ data

It is not a rapid query but it does not take months. ;-)

```{r}
# dataOpenAQ <- NULL
# for (i in 1:length(locationsDelhi)){
#   firstUpdated <- locationsDelhi[i,]$firstUpdated
#   locationURL <- locationsDelhi[i,]$locationURL
#   
#   seqDays <- seq(from=lubridate::ymd(format(firstUpdated, "%Y-%m-%d")),
#                  to=lubridate::ymd("2015-12-31"),
#                  by="1 day")
#   seqDays <- format(seqDays, "%Y-%m-%d")
#   for(i in 1:(length(seqDays)-1)){
#     dataOpenAQTemp <- try(Ropenaq::measurements(location=locationURL,
#                                                      parameter="pm25",
#                                                      limit=1000,
#                                                      date_from=seqDays[i],
#                                                      date_to=seqDays[i+1]), silent=TRUE)
#     print(seqDays[i])
#     
#     if(class(dataOpenAQTemp)[1]!="try-error"){
#       dataOpenAQ <- rbind(dataOpenAQ,
#                                dataOpenAQTemp)
#     }
#     
#   }
#   
# 
# }
# # might be useful later
# dataOpenAQ <- unique(dataOpenAQ)
# save(dataOpenAQ, file="dataOpenAQ.RData")
# write.table(dataOpenAQ, row.names=FALSE, file="dataOpenAQ.csv",
#             sep=",")
load("dataOpenAQ.RData")
```

Put these data in shape.

```{r}
dataOpenAQ <- dplyr::mutate(dataOpenAQ,
                            openAQValue=value,
                            station=location)
dataOpenAQ <- dplyr::select(dataOpenAQ,
                            dateLocal,
                            station,
                            openAQValue)
```

# Comparison

This is the really interesting part I guess.

```{r, warning=FALSE}
for (stationNow in levels(as.factor(dataOpenAQ$station))){
  print(stationNow)
  
  # filter only data for the station
  dataTempCPCB <- dataCPCB[dataCPCB$station==stationNow,]
  dataTempOpenAQ <- dataOpenAQ[dataOpenAQ$station==stationNow,]
  
  # now filter only dates with data from both sources
  minDate <- min(dataTempOpenAQ$dateLocal)
  maxDate <- max(dataCPCB$dateLocal)
  
  dataTempCPCB <- dplyr::filter(dataTempCPCB,
                                dateLocal>=minDate)

  dataTempOpenAQ <- dplyr::filter(dataTempOpenAQ,
                                  dateLocal<=maxDate)
  
  # now combine both data sets
  dataTempCPCB <- dplyr::mutate(dataTempCPCB,
                                sourceData="historic",
                                value=historicValue)%>%
    dplyr::select(dateLocal,
                  value,
                  sourceData)
  dataTempOpenAQ <- dplyr::mutate(dataTempOpenAQ,
                                  sourceData="Open AQ",
                                  value=openAQValue)%>%
    dplyr::select(dateLocal,
                  value,
                  sourceData)
  dataForPlot <- rbind(dataTempCPCB, dataTempOpenAQ)
  
  p <- ggplot() + 
    geom_point(data=dataForPlot,
              aes(x=dateLocal, y=value, col=sourceData))+ 
    ggtitle(stationNow)+ facet_grid(sourceData ~ .)
  print(p)
}


```